### 2020-07-17 latex/sections/background.tex

In this chapter we discuss the algorithms we decided to use, what some other similar 
methods exsist as well as the field in general.

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|p{6cm}|}
    \hline
    \textbf{characteristic} & \textbf{Values} & \textbf{Description}\\
    \hline
    Observable & Fully, Partially & How much of the environment can your agent percieve.\\
    \hline
    Agents & Single, Multi & Are there multiple agents playing in the environment.\\
    \hline
    Deterministic & Deterministic, Stochastic & Do the actions your agent do deterministicly impact the environment\\
    \hline
    Episodic & Sequential, Episodic & Are actions episodic or sequential\\
    \hline
    Static & Static, Semi-Static, Dynamic & ehh  \\
    \hline
    Discrete & Discrete, Continious & Is your environment discrete w.r.t actions (does it end) \\
    \hline
  \end{tabular}
  \caption{Charcteristics of environments}
  \label{tab:env_characteristics}
\end{table}

@@

A different heuristic would be the distance of your most advanced pawn minus your opponents most 
advanced pawn, this heuristic could give you insight into how close you are to winning the game or 
how close your opponent is. As a higher level concept we can call this concept your aggressiveness, 
as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
it is favourable to move your whole team as a unit and play more defensively, so this heuristic 
is probably not one that will lead to the most wins.

@@

Traditionally methods for playing games were searching through the environment using a heuristic to
guide the search. A simple way of doing a heuristic based search would be to give all states a $0$ 
as a score and give states where you win a positive score. We say that those search algorithm 
is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
This method of searching is generally extremenly inefficient as the state-spaces of game environments 
are generally extremely large, and often infinite. For instance an estimate of the state-space 
for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
of the board. The $2^{N*4}$ represent each piece either being alive or captured. So for a small board
$5x4$ the estimated state-space is $1310720$ many states.

@@

More modernly these search methods have been amplified by Machine Learning, in such a way that 
we do not need to figure out a good heuristic for a given state, but rather, we apply a machine 
learning model to learn a function that takes in a state and returns an evaluation of that state.
This can lead to a significant time reduction as we do not need to simulate a whole game from a 
state to recieve its evaluation we rather receive the evaluation from the model.

@@

MCTS is a method of exploring an enviroment in a randomized manner (Monte Carlo is the term implying
randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Backpropagation. They happen sequentially
and repeatedly. MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.

@@

Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
extrapolate some underlying function of that data. This some task like predicting the 
rise and fall of some stock, what the weather will be in a week, and whether an image 
is an image of a dog or a cat. Many datastructures and algorithms are used to achieve this 
goal, but recently the field of ANN/DNN's has been the standard for achieveing the best
results.

@@

Reinforcement Learning is a subfield of Machine Learning where a model learns from experience.
That is, the notion of input/output values changes, the model uses itself to generate input
values by acting in an environment. The environment then returns some result, that could be
losing in a game, making a correct prediction, or any number of outcomes. The result is then
propagated through the model allowing it to improve with this new information.
% When doing moving it's parameters in the direction of the recieved output
% from the environment.

@@

Neural networks (NN) are popular methods within a subfield of ML which is called Deep Learning (DL). 
NN's are created to resemble how the human brain functions. In the brain we have neurons which 
when they get a signal they apply some function to them and if the resulting signal is high 
enough, they fire to the next neuron. This is how it is done in the neural network model as well.
There we have neurons that when they get some input, generally a vector of numbers. The neuron 
takes the sum of that vector, weights the sum by a constant, then applies a non-linear activation 
function to it. The result of doing this is then passed on to the next neuron. Until a final layer 
of neurons is reached. At that point we have a value that the neural network corresponds to the 
input value. This value can be a binary classification (cat or dog image), a regression value (the 
value of a property), or any number of outputs. It can then be said that a neural network is doing 
a function approximation of an input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.

>>>>

In this chapter we discuss the algorithms we decided to use, what some other similar 
methods exist as well as the field in general.

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|p{6cm}|}
    \hline
    \textbf{characteristic} & \textbf{Values} & \textbf{Description}\\
    \hline
    Observable & Fully, Partially & How much of the environment can your agent percieve.\\
    \hline
    Agents & Single, Multi & Are there multiple agents playing in the environment.\\
    \hline
    Deterministic & Deterministic, Stochastic & Do the actions your agent do deterministicly impact the environment\\
    \hline
    Episodic & Sequential, Episodic & Are actions episodic or sequential\\
    \hline
    Static & Static, Semi-Static, Dynamic & ehh  \\
    \hline
    Discrete & Discrete, Continuous & Is your environment discrete w.r.t actions (does it end) \\
    \hline
  \end{tabular}
  \caption{Characteristics of environments}
  \label{tab:env_characteristics}
\end{table}

@@

A different heuristic would be the distance of your most advanced pawn minus your opponents most 
advanced pawn, this heuristic could give you insight into how close you are to winning the game or 
how close your opponent is. As a higher level concept we can call this concept your aggressiveness, 
as it closely resembles how aggressive you are for going in for the win. Generally in Breakthrough 
it is favorable to move your whole team as a unit and play more defensively, so this heuristic 
is probably not one that will lead to the most wins.

@@

Traditionally methods for playing games were searching through the environment using a heuristic to
guide the search. A simple way of doing a heuristic based search would be to give all states a $0$ 
as a score and give states where you win a positive score. We say that those search algorithm 
is not guided, and the algorithm will probably have to evaluate every state in the state-space. 
This method of searching is generally extremely inefficient as the state-spaces of game environments 
are generally extremely large, and often infinite. For instance an estimate of the state-space 
for the game breakthrough is $M * N * 2^{N*4}$ where $M$ is the height of the board $N$ is the width 
of the board. The $2^{N*4}$ represent each piece either being alive or captured. So for a small board
$5x4$ the estimated state-space is $1310720$ many states.

@@

More modernly these search methods have been amplified by Machine Learning, in such a way that 
we do not need to figure out a good heuristic for a given state, but rather, we apply a machine 
learning model to learn a function that takes in a state and returns an evaluation of that state.
This can lead to a significant time reduction as we do not need to simulate a whole game from a 
state to receive its evaluation we rather receive the evaluation from the model.

@@

MCTS is a method of exploring an environment in a randomized manner (Monte Carlo is the term implying
randomness). In MCTS there are four stages. Selection, Expansion, Simulation, and Back-propagation. They happen sequentially
and repeatedly. MCTS is initialized with a tree consisting of the unexpanded initial state of the environment.

@@

Machine Learning (ML) is a research field in which machines apply statistical functions on data to achieve the output that is \textit{correct}, by correct we mean the corresponding result which we expect. Generally this a repetitive process where we look at examples of the data, and the algorithm progressively gets closer to the underlying function of the data it is fed. This process is therefore similar to trial and error for humans. ML is a sub-field of Artificial Intelligence. ML algorithms come in the form of two classes \textit{Classification}, where the algorithm should find a class representing the data its given, and \textit{Regression} where the algorithm should find an underlying continuous numerical function and will return a numerical value representing the input.

Typically ML can be viewed in three different groups, Supervised Learning, Reinforcement Learning (RL), and Unsupervised Learning. Where in Supervised Learning, the algorithm is given data examples and their corresponding outcome. For example, a supervised learning algorithm could be provided with data regarding the weather and the corresponding temperature, the algorithm should then find a pattern within the weather data and find the continuous function represented by the data. This would then be an example of a regression task. Flipping thing example around, if the algorithm would just be provided the temperature and it should tell us whether it is sunny outside or not, that would be a classification task.

In RL the algorithm is given only some input, and then the algorithm tries some outcome generally actions in some environment. Then over an episode~\footnote{a series of outcomes / a timespan} once the episode is finished some reward is given. The algorithm will then learn whether the actions were good actions from the reward. Examples of this are agents playing a game like Flappy Bird, where the data they are given is the state of the game, and they try to either jump or not jump.

%Machine learning (ML) focuses on the using of data and a corresponding outcome w.r.t that data to 
%extrapolate some underlying function of that data. Examples of how we use machine learning is for instance the ability to predicting the 
%rise and fall of some stock, predicting what the weather will be in a week, and whether an image 
%is an image of a dog or a cat. Many data structures and algorithms are used to achieve this 
%goal, but recently the field of ANN/DNN's has been the standard for achieving the best
%results.

%As a general notion we can split machine learning into three different sects those are Supervised Learning, Reinforcement Learning, and Unsupervised Learning.

\subsection{Supervised Learning}

In supervised learning, the data and their corresponding outcomes are already available to us. Examples of the data we can use in supervised learning would be labeled images.

@@

Reinforcement Learning is where a model learns from experience. That is, the notion of input/output values changes, the model uses itself to generate input values by acting in an environment. The environment then returns some result, that could be losing in a game, making a correct prediction, or any number of outcomes. The result is then propagated through the model allowing it to improve with this new information.
% When doing moving it's parameters in the direction of the recieved output
% from the environment.

@@

Neural networks (NN) are popular methods within a sub-field of ML which is called Deep Learning (DL). 
NN's are created to resemble how the human brain functions. In the brain we have neurons which 
when they get a signal they apply some function to them and if the resulting signal is high 
enough, they fire to the next neuron. This is how it is done in the neural network model as well.
There we have neurons that when they get some input, generally a vector of numbers. The neuron 
takes the sum of that vector, weights the sum by a constant, then applies a non-linear activation 
function to it. The result of doing this is then passed on to the next neuron. Until a final layer 
of neurons is reached. At that point we have a value that the neural network corresponds to the 
input value. This value can be a binary classification (cat or dog image), a regression value (the 
value of a property), or any number of outputs. It can then be said that a neural network is doing 
a function approximation of an input to some value. $f_n ( w_n * f_{n-1} (w_{n-1} * \dots f_0(w_0*i))) = o$.

### 2020-07-17 latex/sections/methods.tex

The data collected is $(\pi, \tau, p, v)$, where $\pi$ is the action probabilities provided by 
MCTS for $N$ iterations, $\tau$ is the reward from playing the game, $p$ is the policy vector
from the neural network, and $v$ is the predicted reward of the game by the neural network.

@@

The data collected is back-propagated through the NN moving the weights to the direction of this 
loss function $l = p * \pi + (\tau - v)^2$

>>>>

The data collected is $(\pi, \tau, p, v)$, where $\pi$ is the action probabilities provided by 
MCTS for $N$ iterations, $\tau$ is the reward for the whole episode, $1$ if white wins,
$-1$ if black wins, and $0$ on a tied game, $p$ is the policy vector
from the neural network, and $v$ is the predicted reward of the game by the neural network.

@@

The data collected is backpropagated through the NN moving the weights to the direction of this 
loss function $l = (p * \pi) + (\tau - v)^2$ for each state the NN encountered during self-play.
